Improved state-space models for inference about spatial and temporal variation in abundance from count data

### 2014-06-25 Dail_Mad_ext.tex
Response to reviewer's comments

The negative binomial distribution was strongly supported for ovenbird
initial abundance over the Poisson and zero-inflated Poisson
(Table S3.1A in Appendix 3). %Even compared to the Poisson, there was
%no evidence to support a zero-inflation factor for this species.
The best supported model for $p$ included additive effects of wind speed
and first run of a route by an observer (Table S3.1B, model B.1). First run 
and increasing wind speeds both decreased $p$. All dynamics models 
with immigration were better supported than models without immigration 
(Table S3.1C); the best supported of these was the Ricker-logistic + immigration 
(see Appendix 3, Table S3.2 OR Figure S3.1 for estimates).
%This model estimated $\Lambda$ at 33.4 $\pm$ 4.7, $r$ at 0.026 $\pm$ 0.006, 
%$K$ at  56.9 $\pm$ 9.4, and $\iota$ at 0.283 $\pm$ 0.043.
%The best supported model of those without immigration was the
%Gompertz-logistic, and the geometric-recruitment and the constant model had the least
%support.
%There was little support for an effect of minimum June and July
% temperature on dynamics parameters of the Ricker + Immigration model
%(Table 2A, models A.17 - A.18).
%For both species, 
Constant-recruitment dynamics models estimated unrealisticly high
survival probabilities ($\omega$ = 1 $\pm$ 1.3e-05), whereas % for ovenbird and
%0.935 $\pm$ 0.011 for
%golden-winged warbler). 
the geometric-recruitment and geometric-recruitment + immigration
models estimated nonviably low survival probabilities %for ovenbird 
($\omega$ =
0.058 $\pm$ 0.046 and 0.026 $\pm$ 0.060, respectively). 
%but not for golden-winged warbler ($\omega$ = 0.715 $\pm$ 0.288 for both models).

Estimates for the top ranked model %for ovenbird
%(NB[$\Lambda (.) \alpha (.)$]Ricker-logistic+Immigration[$r(.)K(.) \iota(.)$]p(wind+1st)) 
were similar when
run in the Bayesian framework, except that the estimate of $r$
more than halved 
%% JEFF: How do we explain that????
(Table S3.2 OR Figure S3.1, Appendix 3). %to 0.011 $\pm$ 0.005 [for
%Bayesian model estimates we present mean and SD]). 
When random observer effects
were added, estimates for $\Lambda$ and $K$ increased dramatically
%(to 42.5 $\pm$ 7.3 and 101.8 $\pm$ 21.1,
%respectively),
and the estimate for r was intermediate. The %(0.018 $\pm$ 0.006)
estimate of the intercept for $p$ dropped. % from -1.5 $\pm$
%0.1 to -2.0 $\pm$ 0.1 (on the logit scale), and the estimate of $\sigma_p$ was 0.35 $\pm$ 0.03.  
When environmental stochasticity was added, the estimate for $K$ increased. %(to 112 $\pm$ 58.8) and $r$ was
%again intermediate (0.020 $\pm$ 0.010).  Adding environmental stochasticity alone had little effect
%on estimate of $p$ or $\Lambda$. 
The highest estimates of $\Lambda$ and $K$ came 
from the model that included environmental stochasticity and random observer effects. %(43.8 $\pm$
%7.9 and 141.5 $\pm$ 69.5, respectively).  
Gelman and Rubin diagnostics and visual examination of the chain trajectory and density plots
provided little evidence of lack of convergence for any of the Bayesian models ran.

@@

We thank J. Andrew Royle and T. Scott Sillett for helpful
comments and suggestions. We gratefully acknowledge the U. S. National Park 
Service, the Smithsonian Institution, and [Richard's funding sources] for funding support.

>>>>

The negative binomial distribution was strongly supported for ovenbird
initial abundance over the Poisson and zero-inflated Poisson
(Table S3.1A in Appendix 3). %Even compared to the Poisson, there was
%no evidence to support a zero-inflation factor for this species.
The best supported model for $p$ included additive effects of wind speed
and first run of a route by an observer (Table S3.1B, model B.1). First run 
and increasing wind speeds both decreased $p$. All dynamics models 
with immigration were better supported than models without immigration 
(Table S3.1C); the best supported of these was the Ricker-logistic + immigration 
(see Appendix 3, Fig. S3.1 for estimates).
%This model estimated $\Lambda$ at 33.4 $\pm$ 4.7, $r$ at 0.026 $\pm$ 0.006, 
%$K$ at  56.9 $\pm$ 9.4, and $\iota$ at 0.283 $\pm$ 0.043.
%The best supported model of those without immigration was the
%Gompertz-logistic, and the geometric-recruitment and the constant model had the least
%support.
%There was little support for an effect of minimum June and July
% temperature on dynamics parameters of the Ricker + Immigration model
%(Table 2A, models A.17 - A.18).
%For both species, 
Constant-recruitment dynamics models estimated unrealisticly high
survival probabilities ($\omega$ = 1 $\pm$ 1.3e-05), whereas % for ovenbird and
%0.935 $\pm$ 0.011 for
%golden-winged warbler). 
the geometric-recruitment and geometric-recruitment + immigration
models estimated nonviably low survival probabilities %for ovenbird 
($\omega$ =
0.058 $\pm$ 0.046 and 0.026 $\pm$ 0.060, respectively). 
%but not for golden-winged warbler ($\omega$ = 0.715 $\pm$ 0.288 for both models).

Estimates for the top ranked model %for ovenbird
%(NB[$\Lambda (.) \alpha (.)$]Ricker-logistic+Immigration[$r(.)K(.) \iota(.)$]p(wind+1st)) 
were similar when
run in the Bayesian framework, except that the estimate of $r$
more than halved 
%% JEFF: How do we explain that????
(Appendix 3, Fig. S3.1). %to 0.011 $\pm$ 0.005 [for
%Bayesian model estimates we present mean and SD]). 
When random observer effects
were added, estimates for $\Lambda$ and $K$ increased dramatically
%(to 42.5 $\pm$ 7.3 and 101.8 $\pm$ 21.1,
%respectively),
and the estimate for r was intermediate. The %(0.018 $\pm$ 0.006)
estimate of the intercept for $p$ dropped. % from -1.5 $\pm$
%0.1 to -2.0 $\pm$ 0.1 (on the logit scale), and the estimate of $\sigma_p$ was 0.35 $\pm$ 0.03.  
When environmental stochasticity was added, the estimate for $K$ increased. %(to 112 $\pm$ 58.8) and $r$ was
%again intermediate (0.020 $\pm$ 0.010).  Adding environmental stochasticity alone had little effect
%on estimate of $p$ or $\Lambda$. 
The highest estimates of $\Lambda$ and $K$ came 
from the model that included environmental stochasticity and random observer effects. %(43.8 $\pm$
%7.9 and 141.5 $\pm$ 69.5, respectively).  
Gelman and Rubin diagnostics and visual examination of the chain trajectory and density plots
provided little evidence of lack of convergence for any of the Bayesian models ran.

@@

We thank J. Andrew Royle and T. Scott Sillett for helpful suggestions on our analysis
and manuscript. We gratefully acknowledge the U. S. National Park 
Service, the Smithsonian Institution, and USGS's Breeding Bird Survey program for funding support.
We thank K. Pardieck and D. Zilowkowski for help with data acquisition. J. Sauer 
provided helpful feedback about our analysis.
